---
title: "R_assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data Inspection
Read in data
```{r}
snp <- read.table("snp_position.txt", header = TRUE, fill = TRUE)
fang <- read.table("fang_et_al_genotypes.txt", header = TRUE, fill = TRUE)
```


## Packages
```{r}
if (!require("dplyr")) install.packages("dplyr") 
library(dplyr)

if (!require("ggplot2")) install.packages("ggplot2") 
library(ggplot2)

if (!require("reshape2")) install.packages("reshape2") 
library(reshape2)
```


#### fang
```{r}
typeof(fang)
class(fang)
str(fang)
dim(fang)
```


#### snp_position
```{r}
typeof(snp)
class(snp)
str(snp)
dim(snp)
```


## Data Processing
Sorting Maize from Teosinte files
```{r}
fnames <- unique(fang$Group)
fnames <- as.character(unique(fang$Group))
fznames <- fnames[c(16, 14, 15)]
ftnames <- fnames[c(6, 12, 7)]
zfang <- fang[as.character(fang$Group) %in% fznames, ]
tfang <- fang[as.character(fang$Group) %in% ftnames, ]

```


#### Maize file processing
Save transposed maize file 
```{r}
transz <- t(zfang)
transzdf <- as.data.frame(transz, stringsAsFactors = FALSE)
transzdf$sample_ID <- rownames(transzdf)
names(transzdf) <- as.character(transzdf[1,])
transzdf <-  transzdf[-(1:3),]
```


Merge snp and transposed maize files and reoder columns into SNP ID, Chromosome, postition
```{r}
zmerge <- merge(snp, transzdf, by.x = "SNP_ID", by.y = "Sample_ID")
colzmerge <- zmerge [,c(1,3,4,2,5:1588)]
```


Seperate files by chromosome
```{r}
zChr1 <- colzmerge[colzmerge$Chromosome == "1",]
zChr2 <- colzmerge[colzmerge$Chromosome == "2",]
zChr3 <- colzmerge[colzmerge$Chromosome == "3",]
zChr4 <- colzmerge[colzmerge$Chromosome == "4",]
zChr5 <- colzmerge[colzmerge$Chromosome == "5",]
zChr6 <- colzmerge[colzmerge$Chromosome == "6",]
zChr7 <- colzmerge[colzmerge$Chromosome == "7",]
zChr8 <- colzmerge[colzmerge$Chromosome == "8",]
zChr9 <- colzmerge[colzmerge$Chromosome == "9",] 
zChr10 <- colzmerge[colzmerge$Chromosome == "10",] 
```


Sort position column in each chromosome file as ascending
```{r}
zChr1ascend <- ZChr1[order(as.numeric(as.character(ZChr1$Position))),]
zChr2ascend <- ZChr2[order(as.numeric(as.character(ZChr2$Position))),]
zChr3ascend <- ZChr3[order(as.numeric(as.character(ZChr3$Position))),]
zChr4ascend <- ZChr4[order(as.numeric(as.character(ZChr4$Position))),]
zChr5ascend <- ZChr5[order(as.numeric(as.character(ZChr5$Position))),] 
zChr6ascend <- ZChr6[order(as.numeric(as.character(ZChr6$Position))),] 
zChr7ascend <- ZChr7[order(as.numeric(as.character(ZChr7$Position))),]
zChr8ascend <- ZChr8[order(as.numeric(as.character(ZChr8$Position))),]
zChr9ascend <- ZChr9[order(as.numeric(as.character(ZChr9$Position))),] 
zChr10ascend <- ZChr10[order(as.numeric(as.character(ZChr10$Position))),]
```
  
  
Sort position column in each chromosome file as descending
```{r}
  zChr1descend <- ZChr1[order(as.numeric(as.character(ZChr1$Position)) , decreasing = TRUE),]
  zChr2descend <- ZChr2[order(as.numeric(as.character(ZChr2$Position)) , decreasing = TRUE),]
  zChr3descend <- ZChr3[order(as.numeric(as.character(ZChr3$Position)) , decreasing = TRUE),]
  zChr4descend <- ZChr4[order(as.numeric(as.character(ZChr4$Position)) , decreasing = TRUE),] 
  zChr5descend <- ZChr5[order(as.numeric(as.character(ZChr5$Position)) , decreasing = TRUE),]
  zChr6descend <- ZChr6[order(as.numeric(as.character(ZChr6$Position)) , decreasing = TRUE),]
  zChr7descend <- ZChr7[order(as.numeric(as.character(ZChr7$Position)) , decreasing = TRUE),] 
  zChr8descend <- ZChr8[order(as.numeric(as.character(ZChr8$Position)) , decreasing = TRUE),] 
  zChr9descend <- ZChr9[order(as.numeric(as.character(ZChr9$Position)) , decreasing = TRUE),]
  zChr10descend <- ZChr10[order(as.numeric(as.character(ZChr10$Position)) , decreasing = TRUE),] 
```
 
 
Remove "?" and replace with "-" in descending files
```{r}
  zChr1descend[zChr1 == "?/?"] <- "-/-"
  zChr2descend[zChr2 == "?/?"] <- "-/-"
  zChr3descend[zChr3 == "?/?"] <- "-/-"
  zChr4descend[zChr4 == "?/?"] <- "-/-"
  zChr5descend[zChr5 == "?/?"] <- "-/-"
  zChr6descend[zChr6 == "?/?"] <- "-/-"
  zChr7descend[zChr7 == "?/?"] <- "-/-"
  zChr8descend[zChr8 == "?/?"] <- "-/-"
  zChr9descend[zChr9 == "?/?"] <- "-/-"
  zChr10descend[zChr10 == "?/?"] <- "-/-"
```


Remove "mulptile" SNP position rows by changing "multiple" to NA and then removing the NAs
```{r}
  zChr2ascend[zChr2ascend == "multiple"] <- NA
  zChr2ascend <- zChr2ascend[!is.na(zChr2ascend$Position),]
  zChr4ascend[zChr4ascend == "multiple"] <- NA
  zChr4ascend <- zChr4ascend[!is.na(zChr4ascend$Position),]
  zChr6ascend[zChr6ascend == "multiple"] <- NA
  zChr6ascend <- zChr6ascend[!is.na(zChr6ascend$Position),]
  zChr7ascend[zChr7ascend == "multiple"] <- NA
  zChr7ascend <- zChr7ascend[!is.na(zChr7ascend$Position),]
  zChr9ascend[zChr9ascend == "multiple"] <- NA
  zChr9ascend <- zChr9ascend[!is.na(zChr9ascend$Position),]
  zChr2descend[zChr2descend == "multiple"] <- NA
  zChr2descend <- zChr2descend[!is.na(zChr2descend$Position),]
  zChr4descend[zChr4descend == "multiple"] <- NA
  zChr4descend <- zChr4descend[!is.na(zChr4descend$Position),]
  zChr6descend[zChr6descend == "multiple"] <- NA
  zChr6descend <- zChr6descend[!is.na(zChr6descend$Position),]
  zChr7descend[zChr7descend == "multiple"] <- NA
  zChr7descend <- zChr7descend[!is.na(zChr7descend$Position),]
  zChr9descend[zChr9descend == "multiple"] <- NA
  zChr9descend <- zChr9descend[!is.na(zChr9descend$Position),]
```


Remove columns that are not SNP ID, chromosome, position, and genotype columns
```{r}
zChr1ascend <- zChr1ascend[, -c(4:15)]
zChr2ascend <- zChr2ascend[, -c(4:15)]
zChr3ascend <- zChr3ascend[, -c(4:15)]
zChr4ascend <- zChr4ascend[, -c(4:15)]
zChr5ascend <- zChr5ascend[, -c(4:15)]
zChr6ascend <- zChr6ascend[, -c(4:15)]
zChr7ascend <- zChr7ascend[, -c(4:15)]
zChr8ascend <- zChr8ascend[, -c(4:15)]
zChr9ascend <- zChr9ascend[, -c(4:15)]
zChr10ascend <- zChr10ascend[, -c(4:15)]
zChr1descend <- zChr1descend[, -c(4:15)]
zChr2descend <- zChr2descend[, -c(4:15)]
zChr3descend <- zChr3descend[, -c(4:15)]
zChr4descend <- zChr4descend[, -c(4:15)]
zChr5descend <- zChr5descend[, -c(4:15)]
zChr6descend <- zChr6descend[, -c(4:15)]
zChr7descend <- zChr7descend[, -c(4:15)]
zChr8descend <- zChr8descend[, -c(4:15)]
zChr9descend <- zChr9descend[, -c(4:15)]
zChr10descend <- zChr10descend[, -c(4:15)]
```


Create text files for each chromosome with ascending and descending SNP positions
```{r}
write.table(zChr1ascend, "Chromosome1_Maize_ascending.txt", sep = "\t")
write.table(zChr2ascend, "Chromosome2_Maize_ascending.txt", sep = "\t")
write.table(zChr3ascend, "Chromosome3_Maize_ascending.txt", sep = "\t")
write.table(zChr4ascend, "Chromosome4_Maize_ascending.txt", sep = "\t")
write.table(zChr5ascend, "Chromosome5_Maize_ascending.txt", sep = "\t")
write.table(zChr6ascend, "Chromosome6_Maize_ascending.txt", sep = "\t")
write.table(zChr7ascend, "Chromosome7_Maize_ascending.txt", sep = "\t")
write.table(zChr8ascend, "Chromosome8_Maize_ascending.txt", sep = "\t")
write.table(zChr9ascend, "Chromosome9_Maize_ascending.txt", sep = "\t")
write.table(zChr10ascend, "Chromosome10_Maize_ascending.txt", sep = "\t")
write.table(zChr1descend, "Chromosome1_Maize_descending.txt", sep = "\t")
write.table(zChr2descend, "Chromosome2_Maize_descending.txt", sep = "\t")
write.table(zChr3descend, "Chromosome3_Maize_descending.txt", sep = "\t")
write.table(zChr4descend, "Chromosome4_Maize_descending.txt", sep = "\t")
write.table(zChr5descend, "Chromosome5_Maize_descending.txt", sep = "\t")
write.table(zChr6descend, "Chromosome6_Maize_descending.txt", sep = "\t")
write.table(zChr7descend, "Chromosome7_Maize_descending.txt", sep = "\t")
write.table(zChr8descend, "Chromosome8_Maize_descending.txt", sep = "\t")
write.table(zChr9descend, "Chromosome9_Maize_descending.txt", sep = "\t")
write.table(zChr10descend, "Chromosome10_Maize_descending.txt", sep = "\t")
```


#### Teosinte file processing
Transpose teosinte file 
```{r}
  transt <- t(tfang)
  transtdf <- as.data.frame(transt, stringsAsFactors = FALSE)
  transtdf$sample_ID <- rownames(transtdf)
  names(transtdf) <- as.character(transtdf[1,])
  transtdf <-  transtdf[-(1:3),]
```


Merge snp file to teosinte file
```{r}
  tmerge <- merge(snp, transtdf, by.x = "SNP_ID", by.y = "Sample_ID")
  coltmerge <- tmerge [,c(1,3,4,2,5:990)]
```


Seperate files by chromosome
```{r}
  tChr1 <- coltmerge[coltmerge$Chromosome == "1",]
  tChr2 <- coltmerge[coltmerge$Chromosome == "2",]
  tChr3 <- coltmerge[coltmerge$Chromosome == "3",]
  tChr4 <- coltmerge[coltmerge$Chromosome == "4",]
  tChr5 <- coltmerge[coltmerge$Chromosome == "5",]
  tChr6 <- coltmerge[coltmerge$Chromosome == "6",]
  tChr7 <- coltmerge[coltmerge$Chromosome == "7",]
  tChr8 <- coltmerge[coltmerge$Chromosome == "8",]
  tChr9 <- coltmerge[coltmerge$Chromosome == "9",]  
  tChr10 <- coltmerge[coltmerge$Chromosome == "10",]
```


Sort position column in each chromosome file as ascending
```{r}
tChr1ascend <- tChr1[order(as.numeric(as.character(tChr1$Position))),]
tChr2ascend <- tChr2[order(as.numeric(as.character(tChr2$Position))),]
tChr3ascend <- tChr3[order(as.numeric(as.character(tChr3$Position))),]
tChr4ascend <- tChr4[order(as.numeric(as.character(tChr4$Position))),]  
tChr5ascend <- tChr5[order(as.numeric(as.character(tChr5$Position))),]  
tChr6ascend <- tChr6[order(as.numeric(as.character(tChr6$Position))),]  
tChr7ascend <- tChr7[order(as.numeric(as.character(tChr7$Position))),]  
tChr8ascend <- tChr8[order(as.numeric(as.character(tChr8$Position))),]  
tChr9ascend <- tChr9[order(as.numeric(as.character(tChr9$Position))),] 
tChr10ascend <- tChr10[order(as.numeric(as.character(tChr10$Position))),] 
```


Sort position column in each chromosome file as descending
```{r}
tChr1descend <- tChr1[order(as.numeric(as.character(tChr1$Position)) , decreasing = TRUE),]
tChr2descend <- tChr2[order(as.numeric(as.character(tChr2$Position)) , decreasing = TRUE),]
tChr3descend <- tChr3[order(as.numeric(as.character(tChr3$Position)) , decreasing = TRUE),]
tChr4descend <- tChr4[order(as.numeric(as.character(tChr4$Position)) , decreasing = TRUE),] 
tChr5descend <- tChr5[order(as.numeric(as.character(tChr5$Position)) , decreasing = TRUE),]
tChr6descend <- tChr6[order(as.numeric(as.character(tChr6$Position)) , decreasing = TRUE),]
tChr7descend <- tChr7[order(as.numeric(as.character(tChr7$Position)) , decreasing = TRUE),] 
tChr8descend <- tChr8[order(as.numeric(as.character(tChr8$Position)) , decreasing = TRUE),] 
tChr9descend <- tChr9[order(as.numeric(as.character(tChr9$Position)) , decreasing = TRUE),]
tChr10descend <- tChr10[order(as.numeric(as.character(tChr10$Position)) , decreasing = TRUE),] 
```


Remove "?" and replace with "-" in descending files
```{r}
tChr1descend[tChr1 == "?/?"] <- "-/-"
tChr2descend[tChr2 == "?/?"] <- "-/-"
tChr3descend[tChr3 == "?/?"] <- "-/-"
tChr4descend[tChr4 == "?/?"] <- "-/-"
tChr5descend[tChr5 == "?/?"] <- "-/-"
tChr6descend[tChr6 == "?/?"] <- "-/-"
tChr7descend[tChr7 == "?/?"] <- "-/-"
tChr8descend[tChr8 == "?/?"] <- "-/-"
```

  
Remove "mulptile" SNP position rows by changing "multiple" to NA and then removing the NAs
```{r}
tChr2ascend[tChr2ascend == "multiple"] <- NA
tChr2ascend <- tChr2ascend[!is.na(tChr2ascend$Position),]
tChr4ascend[tChr4ascend == "multiple"] <- NA
tChr4ascend <- tChr4ascend[!is.na(tChr4ascend$Position),]
tChr7ascend[tChr7ascend == "multiple"] <- NA
tChr7ascend <- tChr7ascend[!is.na(tChr7ascend$Position),]
tChr9ascend[tChr9ascend == "multiple"] <- NA
tChr9ascend <- tChr9ascend[!is.na(tChr9ascend$Position),]
tChr2descend[tChr2descend == "multiple"] <- NA
tChr2descend <- tChr2descend[!is.na(tChr2descend$Position),]
tChr2descend <- tChr2descend[!is.na(tChr2descend$Position),]
tChr4descend[tChr4descend == "multiple"] <- NA
tChr4descend <- tChr4descend[!is.na(tChr4descend$Position),]
tChr7descend[tChr7descend == "multiple"] <- NA
tChr7descend <- tChr7descend[!is.na(tChr7descend$Position),]
tChr9descend[tChr9descend == "multiple"] <- NA
tChr9descend <- tChr9descend[!is.na(tChr9descend$Position),]
```


Remove columns that are not SNP ID, chromosome, position, and genotype columns
```{r}
tChr1ascend <- tChr1ascend[, -c(4:15)]
tChr2ascend <- tChr2ascend[, -c(4:15)]
tChr3ascend <- tChr3ascend[, -c(4:15)]
tChr4ascend <- tChr4ascend[, -c(4:15)]
tChr5ascend <- tChr5ascend[, -c(4:15)]
tChr6ascend <- tChr6ascend[, -c(4:15)]
tChr7ascend <- tChr7ascend[, -c(4:15)]
tChr8ascend <- tChr8ascend[, -c(4:15)]
tChr9ascend <- tChr9ascend[, -c(4:15)]
tChr10ascend <- tChr10ascend[, -c(4:15)]
tChr1descend <- tChr1descend[, -c(4:15)]
tChr2descend <- tChr2descend[, -c(4:15)]
tChr3descend <- tChr3descend[, -c(4:15)]
tChr4descend <- tChr4descend[, -c(4:15)]
tChr5descend <- tChr5descend[, -c(4:15)]
tChr6descend <- tChr6descend[, -c(4:15)]
tChr7descend <- tChr7descend[, -c(4:15)]
tChr8descend <- tChr8descend[, -c(4:15)]
tChr9descend <- tChr9descend[, -c(4:15)]
tChr10descend <- tChr10descend[, -c(4:15)]
```


Create text files for each chromosome with ascending and descending SNP positions
```{r}
write.table(tChr1ascend, "Chromosome1_Teosinte_ascending.txt", sep = "\t")
write.table(tChr2ascend, "Chromosome2_Teosinte_ascending.txt", sep = "\t")
write.table(tChr3ascend, "Chromosome3_Teosinte_ascending.txt", sep = "\t")
write.table(tChr4ascend, "Chromosome4_Teosinte_ascending.txt", sep = "\t")
write.table(tChr5ascend, "Chromosome5_Teosinte_ascending.txt", sep = "\t")
write.table(tChr6ascend, "Chromosome6_Teosinte_ascending.txt", sep = "\t")
write.table(tChr7ascend, "Chromosome7_Teosinte_ascending.txt", sep = "\t")
write.table(tChr8ascend, "Chromosome8_Teosinte_ascending.txt", sep = "\t")
write.table(tChr9ascend, "Chromosome9_Teosinte_ascending.txt", sep = "\t")
write.table(tChr10ascend, "Chromosome10_Teosinte_ascending.txt", sep = "\t")
write.table(tChr1descend, "Chromosome1_Teosinte_descending.txt", sep = "\t")
write.table(tChr2descend, "Chromosome2_Teosinte_descending.txt", sep = "\t")
write.table(tChr3descend, "Chromosome3_Teosinte_descending.txt", sep = "\t")
write.table(tChr4descend, "Chromosome4_Teosinte_descending.txt", sep = "\t")
write.table(tChr5descend, "Chromosome5_Teosinte_descending.txt", sep = "\t")
write.table(tChr6descend, "Chromosome6_Teosinte_descending.txt", sep = "\t")
write.table(tChr7descend, "Chromosome7_Teosinte_descending.txt", sep = "\t")
write.table(tChr8descend, "Chromosome8_Teosinte_descending.txt", sep = "\t")
write.table(tChr9descend, "Chromosome9_Teosinte_descending.txt", sep = "\t")
write.table(tChr10descend, "Chromosome10_Teosinte_descending.txt", sep = "\t")
```


## Data Analysis

Load files
```{r}
snp <- read.table("snp_position.txt", header = TRUE, stringsAsFactors = FALSE, sep = "\t")
fang <- read.table("fang_et_al_genotypes.txt", header = TRUE, stringsAsFactors = FALSE, na.strings = "?/?")

```


Check the structure 
```{r}
str(snp)
str(fang)
```


Transposing the fang dataset
```{r}
tfang <- as.data.frame(t(fang), stringsAsFactors = FALSE)
tfang$SampleID <- row.names(tfang)
str(tfang$SampleID)
str(tfang)
```


Merging snp and tfang datasets
```{r}
mdata <- merge(snp,tfang,by.x = "SNP_ID", by.y = "SampleID", all.y = TRUE)

View(mdata)
```


Plot by number of snps per chromosome
```{r}

ggplot(mdata) + geom_bar(aes(x=Chromosome))
```


Number of SNPS per group
```{r}
ggplot(fang)+geom_bar(aes(fang$Group))+xlab("SNPS Group")+ylab("Total SNPS")
```


Melt and format data 
```{r}
fangrm <- fang[,(-2)]
meltfang <- melt(fangrm, id.vars = c("Sample_ID", "Group"))
```


Replace values for homozygous and heterzygous
```{r}
meltfang[meltfang == "A/A" | meltfang == "C/C" | meltfang == "G/G" | meltfang == "T/T"] <- 0

meltfang[meltfang == "C/G" | meltfang == "C/A" | meltfang == "C/T" | meltfang == "G/T" | meltfang == "G/A" | meltfang == "A/T" | meltfang == "A/C" | meltfang == "A/G"] <- 1
```


Replace groups with species names
```{r}
group <- meltfang$Group

meltfang$Species_ID <-group

meltfang[meltfang == "ZMMIL" | meltfang == "ZMMLR" | meltfang == "ZMMMR"] <- "maize"

meltfang[meltfang == "ZMPBA" | meltfang == "ZMPIL" | meltfang == "ZMPJA"] <- "teosinte"

unique(meltfang$Species_ID)

meltfang[meltfang == "TRIPS" | meltfang == "ZDIPL" | meltfang == "ZPERR" | meltfang == "ZLUXR" | meltfang == "ZMHUE" | meltfang == "ZMXCH" | meltfang == "ZMXCP" | meltfang == "ZMXNO" | meltfang == "ZMXNT" | meltfang == "ZMXIL"] <- "tripsacum"

meltfang$Group <- group

meltfang[meltfang == "?/?"] <- NA
```


Plot homozygous and heterzygous proportions
```{r}
ggplot(meltfang, aes(x = Group, fill = value)) + geom_bar(position = "fill") + scale_fill_discrete(name="Heterzygousity Proportions", breaks=c("0", "1", "NA"), labels = c("Homozygous", "Heterozygous", "NA")) + facet_grid(Species_ID ~ .)
```


Plot positions on each chromosome
```{r}
meltsnp <- melt(snp, id.vars = "Chromosome", "Position")
meltsnp <- meltsnp[order(meltsnp$Chromosome),]
meltsnp <- meltsnp[-c(1:2),]
meltsnp <- meltsnp[-c(950:1015),]
meltsnp <- meltsnp[order(meltsnp$value),]
meltsnp <- meltsnp[-c(1),]
meltsnp <- meltsnp[-c(940:950),]
ggplot(meltsnp, aes(x=value, y=Chromosome)) + scale_x_discrete(breaks = seq(10000, 1000000000, by = 10000)) + geom_point() + xlab("Position") + ylab("Chromosome")
```




